Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-05-11 Manuscript/PLoSONEresponse/response.tex

%We feel that the reviewer has slightly misunderstood the term ``statistics'' in this case. Firstly, p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e. significant ones) when the sample sizes are big; (2) concerning the confidence intervals, we actually did systematically calculate them; ans (3) we are not convinced by a Bayesian approach for comparing distributions.

%-------------

\textcolor{blue}{B-The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}

The reviewer is (we believe) referring to their comment below (number 13). We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only and we are ignoring the branch length. The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of the moves of the MCMC sampler for the three priors combined). Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.4 CPU centuries. Note that the use of a starting tree that is not a prior and just ``the ``true" tree's 226 topology as a starting tree'' (line 226) is not a prior and was used to avoid the chains to be stuck in local optima.

@@

Both birth ($\lambda$) and death ($\mu$) parameters are probabilities bounded between 0 and 1. We have added this to the new version of the manuscript:

%-------------

\textcolor{blue}{5-line 118: with 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}

Yes. We have clarified as follows in our revision (lines @118 to @119):

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
[We] select only trees with 25 living and 25 fossil taxa. The fossil taxa were considered as unique tips at the end of extinct lineages.
\end{minipage}

@@

See our response to comment C above.

%-------------

\textcolor{blue}{9-line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking. It makes a difference. If you mask out $M_F$ percentage of cells, and then mask $M_C$ percentage of columns, it is possible (indeed very probable) that the percentage of missing data in fossil characters that remain will not be $M_F$ after some of the morphological columns are removed. If you mask by $M_C$ first, and then $M_F$, the resulting matrices will always have $M_F$ percentage missing data (modulo rounding error) in the simulated realizations.}

We clarified this point ($M_C$ comes second) in the new version of the manuscript (lines @173 to @183). We also, clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%) in the same revised paragraph (lines @173 to @183).

@@

Our simulations were designed to test the effects of missing data in as realistic a manner as possible given our constraints. Thus, we do not feel that this additional analysis would add much to the paper, especially since it is already rather long and complex.

%-------------

\textcolor{blue}{12-Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}

We implemented this step of keeping trees with a minimum total median bootstrap to make the topological comparisons more robust. If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the low supported nodes. In fact, because we repeat the tree inference 125 times ($5^3$ parameters combinations), it is likely to have 125 different pseudo-random solution for the nodes with really low phylogentic support and that, therefore, the observed changes in topologies are just linked to the pseudo-random replication. Note however that we did not exculde this scenario from our protocol the trees selected with a median bootstrap $>$ 50 could well contain some unresolved nodes with a low bootstrap ($<$ 50). We've added the following justification in our revision (lines @215 to @220):

@@

See our response to comment B above.

%-------------

\textcolor{blue}{14-line 254: RF measures the difference between the number of clades and the twice number of shared clades across two trees. The statement that it ``measures the number of shared clades across two trees" should be revised.}

We revised this statement lines @267 to @269 in our revision:

@@

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
The Robinson-Foulds distance, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees.
\end{minipage}

%-------------

\textcolor{blue}{15-line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}

We thanks the reviewer for pointing at this error and we changed the following sentence (lines @271 to @273 in our revision):

@@

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
\end{minipage}

%-------------

\textcolor{blue}{16-The ``scaling" to produce the Normalized RF ``distance" is confusing. It is also inverting the sign of the distance, so it has become a measure of similarity.}

We agree that this is confusing and thank the reviewer for pointing this inconsistency out. We changed the mentions of ``Normalised Robinson-Foulds distance'' and ``Normalised Triplets distance'' to ``Normalised Robinson-Foulds metric'' and ``Normalised Triplets metric'' throughout the our revision. Additionally, we clarified the scaling method as follows (lines @276 to @284 in our revision):

@@

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
It is bounded between zero when the two trees are identical and $\binom{n}{3}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
\end{minipage}

%-------------

\textcolor{blue}{18-Using the BC stat may be similar in spirit to using a t-test, but it is not equivalent.}

We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``similar" ; lines @322 to @324 in our revision):

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
\end{minipage}

@@

\textcolor{blue}{21-line 398: ``mMthods" $->$ ``Methods"}

We thanks the reviewer for pointing out this typo and fixed it.

\textcolor{blue}{22-line 420: contra the statement here, a consensus collapsing to a polytomy will affect the RF distance (could go up or down, but it will change).}

We revised this statement at lines @434 to @442 in our revision:

@@

\textcolor{blue}{25-The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees. I'm not sure that it makes much difference, but it is odd.}

We followed the protocol in the cited reference (Bogdanowicz 2012) where the normalisation of the metrics is done on the mean value of that metric when comparing to random trees. Using random Yule trees is the best way to generate purely random trees since they are not based on any evolutionary scenario and takes only one parameter which is the number of tips (e.g. as opposed to the Birth-Death tree that assumes more realistic evolutionary scenario and takes multiple parameters).
%-------------

\textcolor{blue}{26-The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0. This is not true of either distance metric here. So the real range of values is (mean-max)/mean for both the RF and triples distance.}

As clarified in our revision (lines @276 to @284), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.
%-------------

### 2015-05-12 Manuscript/PLoSONEresponse/OnlineAppendices-LaTeXSuppFiles/SupportingInformation3.tex
Updated final response

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryFigures/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
    \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey). The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.} 
\label{Fig_Supp_BootBayt_allparam} 
\end{figure}

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
    \caption{The effects of missing data on topological recovery using Maximum Likelihood trees. The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient. The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
\label{Fig_Supp_paircomp_ML}
\end{figure}

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
    \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees. The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient. The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
\label{Fig_Supp_paircomp_Boot}
\end{figure}

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PPairwiseComp-Bayt-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
    \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution. The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient. TThe Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
\label{Fig_Supp_paircomp_Bayt}
\end{figure}

%Summary of metrics values for all parameters combinations
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
\label{Tab_Supp_summary_metric_allparam}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
                                        & $Tr$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
                                        & $Tr$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
                                        & $Tr$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
                                        & $Tr$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the ML parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
                                        & $Tr$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
  Bayesian consensus                    & $RF$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
                                        & $Tr$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
                                        & $Tr$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
                                        & $Tr$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MF parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{F}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MF}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
                                        & $Tr$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
                                        & $Tr$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
                                        & $Tr$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MC parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{C}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MC}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
                                        & $Tr$ & 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
                                        & $Tr$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
                                        & $Tr$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of the BC between pairs of methods for the ML parameter
\begin{landscape}
\begin{table}[ht]
\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
\label{Tab_Supp_summary_BC_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
                                                                       & $Tr$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
                                                                       & $Tr$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
                                                                       & $Tr$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
                                                                       & $Tr$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
                                                                       & $Tr$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
                                                                       & $Tr$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

>>>>

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryFigures/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
    \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey). The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.} 
\label{Fig_Supp_BootBayt_allparam} 
\end{figure}

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
    \caption{The effects of missing data on topological recovery using Maximum Likelihood trees. The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient. The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
\label{Fig_Supp_paircomp_ML}
\end{figure}

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
    \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees. The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient. The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
\label{Fig_Supp_paircomp_Boot}
\end{figure}

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PPairwiseComp-Bayt-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
    \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution. The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient. TThe Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
\label{Fig_Supp_paircomp_Bayt}
\end{figure}

%Summary of metrics values for all parameters combinations
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
\label{Tab_Supp_summary_metric_allparam}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
                                        & $Tr$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
                                        & $Tr$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
                                        & $Tr$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
                                        & $Tr$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the ML parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
                                        & $Tr$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
  Bayesian consensus                    & $RF$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
                                        & $Tr$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
                                        & $Tr$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
                                        & $Tr$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MF parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MF}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
                                        & $Tr$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
                                        & $Tr$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
                                        & $Tr$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MC parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MC}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
                                        & $Tr$ & 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
                                        & $Tr$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
                                        & $Tr$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of the BC between pairs of methods for the ML parameter
\begin{landscape}
\begin{table}[ht]
\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
\label{Tab_Supp_summary_BC_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
                                                                       & $Tr$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
                                                                       & $Tr$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
                                                                       & $Tr$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
                                                                       & $Tr$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
                                                                       & $Tr$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
                                                                       & $Tr$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}
